Spending too much time at the Galactic bar: chaotic fanning of the Ophiuchus stream

### 2016-04-17 ms.tex
shorten abstract

\begin{abstract}
The Ophiuchus stellar stream is peculiar: (1) its length is short given the age of its constituent stars, and (2) several probable member stars that lie close in both sky position and velocity have dispersions in these dimensions that far exceed those seen within the stream.
The stream's proximity to the Galactic center suggests that the bar must have a significant influence on its dynamical history: The triaxiality and time-dependence of the bar may generate chaotic orbits in the vicinity of the stream that can greatly affect its morphology.
We explore this hypothesis with models of stream formation along orbits consistent with Ophiuchus' properties in a Milky Way potential model that includes a rotating bar.
We find that in all choices for the rotation parameters of the bar, orbits fit to the stream are strongly chaotic.
Mock streams generated along these orbits qualitatively match the observed properties of the stream: because of chaos, stars stripped early generally form low-density, high-dispersion ``fans'' leaving only the most recently disrupted material detectable as a strong over-density.
Our models predict that there should be more low-surface-brightness tidal debris than detected so far, likely with a complex phase-space morphology.
The existence of or lack of these features around the Ophiuchus stream would provide an interesting constraint on the properties of the Milky Way bar and would help distinguish between formation scenarios for the stream.
This is the first time that chaos has been used to explain the properties of a stellar stream and is the first demonstration of the dynamical importance of chaos in the Galactic halo.
The existence of long, thin streams around the Milky Way---presumably formed along non- or weakly-chaotic orbits---may represent only a subset of the total population of disrupted satellites.
\end{abstract}

>>>>

\begin{abstract}
The Ophiuchus stellar stream is peculiar: (1) its length is short given the age of its constituent stars, and (2) several probable member stars have dispersions in sky position and velocity that far exceed those seen within the stream. The stream's proximity to the Galactic center suggests that its dynamical history is significantly influenced by the Galactic bar. We explore this hypothesis with models of stream formation along orbits consistent with Ophiuchus' properties in a Milky Way potential model that includes a rotating bar. In all choices for the rotation parameters of the bar, orbits fit to the stream are strongly chaotic. Mock streams generated along these orbits qualitatively match the observed properties of the stream: because of chaos, stars stripped early generally form low-density, high-dispersion ``fans'' leaving only the most recently disrupted material detectable as a strong over-density. Our models predict that there should be a significant amount of low-surface-brightness tidal debris around the stream with a complex phase-space morphology. The existence of or lack of these features could provide interesting constraints on the Milky Way bar and would rule out formation scenarios for the stream. This is the first time that chaos has been used to explain the properties of a stellar stream and is the first demonstration of the dynamical importance of chaos in the Galactic halo. The existence of long, thin streams around the Milky Way, presumably formed along non- or weakly-chaotic orbits, may represent only a subset of the total population of disrupted satellites.
\end{abstract}
